---
title: "UDBio16S"
author: "JReceveur"
date: "November 2, 2017"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=14, fig.height=10)
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
```

#Import
``` {r import, include=FALSE, messages=FALSE, warnings=FALSE,echo= FALSE}
library(vegan)
library(MASS)
library(ggplot2)
library(plyr)
library(dplyr)
library(magrittr)
#library(mctoolsr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(randomForest)
library(knitr)
library(ape)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#000000","#CC79A7")
theme_set(theme_bw(base_size = 15)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()))
biom=import_biom("C:\\Users\\Joe Receveur\\Documents\\MSU data\\UDBiofilms\\UDBF_meta_ns_no_mito.biom",parseFunction= parse_taxonomy_greengenes)
physeq=biom
sample_data(physeq)$Date = factor(sample_data(physeq)$Date, levels = c("17-May-16","31-May-16","14-Jul-16","28-Jul-16","16-Sep-16","30-Sep-16","16-Nov-16","30-Nov-16")) 

sample_data(physeq)$Treatment2=c("Modified","Reference","Modified","Modified","Modified","Reference","Modified","Modified","Reference","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Reference","Reference","Reference","Modified","Reference","Modified","Modified","Modified","Modified","Modified","Modified","Reference","Modified","Reference","Modified","Modified","Reference","Modified","Reference","Reference","Modified","Reference","Modified","Reference","Reference","Modified","Modified","Reference","Modified","Modified","Reference","Reference","Reference","Modified","Reference","Modified","Modified","Modified","Modified","Modified","Reference","Modified","Reference","Reference","Reference","Modified","Modified","Reference","Modified","Reference","Modified","Reference","Modified","Reference","Reference","Modified","Modified","Modified","Reference","Modified","Modified","Modified","Modified","Modified","Modified","Reference","Modified","Reference","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Modified","Reference","Reference","Modified","Modified","Reference","Modified","Modified","Modified","Modified","Modified","Reference","Modified","Reference","Reference","Reference","Reference","Modified","Modified","Modified","Modified","Reference","Reference")
sample_data(physeq)$DateTreat= paste0(sample_data(physeq)$Treatment2," ",sample_data(physeq)$Date) #creates DateTreat variable

#fixes x-axis labels

```
Data Location
```{r Data Location, echo=FALSE}
"C:\\Users\\Joe Receveur\\Documents\\MSU data\\UDBiofilms\\UDBF_meta_ns_no_mito.biom"
```
#Alpha Diversity
##Observed Species

```{r,echo=FALSE}
plot_richness(physeq, x="Date",color="Date", measures=c("Observed"),)+geom_boxplot(aes(x=Date, y=value, color=Date), alpha=0.05)+facet_wrap(~Treatment)+ylab("Observed Species")

```

##Simpson Diversity
```{r,echo=FALSE}

plot_richness(physeq, x="Date",color="Date", measures=c("Simpson"),)+geom_boxplot(aes(x=Date, y=value, color=Date), alpha=0.05)+facet_wrap(~Treatment2)+ylab("Simpson Diversity")

```

##Species Evenness
```{r, echo=FALSE}
Evenness=read.table("C:\\Users\\Joe Receveur\\Documents\\MSU data\\UDBiofilms\\UDBFevennessPielou.txt", header=TRUE)
Evenness$Date = factor(Evenness$Date, levels = c("17-May-16","31-May-16","14-Jul-16","28-Jul-16","16-Sep-16","30-Sep-16","16-Nov-16","30-Nov-16")) #fixes x-axis labels
ggplot(Evenness, aes(x=Date, y=Evenness,color=Date))+facet_wrap(~Treatment)+geom_point()+
  geom_boxplot(aes(x=Date, y=Evenness, color=Date))+ylab("Species Evenness")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


#Taxa Plots
```{r filteringForFamily2 ,echo=FALSE}
GPr  = transform_sample_counts(physeq, function(x) x / sum(x) ) #transform samples based on relative abundance
GPrPhylum=tax_glom(GPr, "Phylum")
PhylumLevel = filter_taxa(GPrPhylum, function(x) mean(x) > 1e-3, TRUE) #filter out any taxa lower tha 0.1%
GPrFamily=tax_glom(GPr,"Family")
FamilyLevel = filter_taxa(GPrFamily, function(x) mean(x) > 1e-2, TRUE) #filter out any taxa lower tha 1%

```

##Phylum Level
```{r PhylumRA, echo=FALSE, warning=FALSE}
TrtDate=merge_samples(PhylumLevel,"Trt_Date")
#sample_data(TrtDate)
sample_data(TrtDate)$Treatment=c("Heavy","Heavy","Heavy","Heavy","Heavy","Heavy","Heavy","Heavy","Moderate","Moderate","Moderate","Moderate","Moderate","Moderate","Moderate","Moderate","Reference","Reference","Reference","Reference","Reference","Reference","Reference","Reference")
sample_data(TrtDate)$Date=c("11/16","11/30","5/17","5/31","7/14","7/28","9/16","9/30","11/16","11/30","5/17","5/31","7/14","7/28","9/16","9/30","11/16","11/30","5/17","5/31","7/14","7/28","9/16","9/30")
sample_data(TrtDate)$Date = factor(sample_data(TrtDate)$Date, levels = c("5/17", "5/31","7/14","7/28","9/16","9/30","11/16","11/30")) #fixes x-axis labels

sample_data(TrtDate)$Instar=sample_names(TrtDate)
#sample_data(Instar)
TrtDate=transform_sample_counts(TrtDate, function(x) 100*x/sum(x)) #merging samples #(averaging)
TrtDatePhylumplot=plot_bar(TrtDate, "Date","Abundance", fill='Phylum')  +ylab("Relative Bacterial Abundance (> 0.1%)")+facet_grid(Treatment ~ .)#+scale_fill_manual(values=cbPalette)
TrtDatePhylumplot+ theme(axis.text.x = element_text(angle = 45, hjust = 1))#+ theme(legend.justification=c(0.05,0.95), legend.position=c(0.05,0.95))

```

##Family Level Relative Abundance
```{r Summarizing2, echo=FALSE,warning=FALSE}
df <- psmelt(FamilyLevel)
df$Abundance=df$Abundance*100
Trtdata <- ddply(df, c("Family", "Treatment"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
```
```{r TreatmentPlot2, echo=FALSE}
Plot=ggplot(Trtdata, aes(x=Treatment, y=mean,fill=Treatment))+facet_wrap(~Family)
Plot+geom_bar(stat="identity")+ geom_errorbar(aes(ymin=mean-se, ymax=mean+se),col="black")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ylab("Relative Abundance (> 1%)")
```

##Family level date by treatment 
```{r RADatexTrtFamily2, echo=FALSE, warning=FALSE, error=FALSE}
TrtDate=merge_samples(FamilyLevel,"Trt_Date")
#sample_data(TrtDate)
sample_data(TrtDate)$Treatment=c("Heavy","Heavy","Heavy","Heavy","Heavy","Heavy","Heavy","Heavy","Moderate","Moderate","Moderate","Moderate","Moderate","Moderate","Moderate","Moderate","Reference","Reference","Reference","Reference","Reference","Reference","Reference","Reference")
sample_data(TrtDate)$Date=c("11/16","11/30","5/17","5/31","7/14","7/28","9/16","9/30","11/16","11/30","5/17","5/31","7/14","7/28","9/16","9/30","11/16","11/30","5/17","5/31","7/14","7/28","9/16","9/30")
sample_data(TrtDate)$Date = factor(sample_data(TrtDate)$Date, levels = c("5/17", "5/31","7/14","7/28","9/16","9/30","11/16","11/30")) #fixes x-axis labels

sample_data(TrtDate)$Instar=sample_names(TrtDate)
#sample_data(Instar)
TrtDate=transform_sample_counts(TrtDate, function(x) 100*x/sum(x)) #merging samples #(averaging)
TrtDateFamilyplot=plot_bar(TrtDate, "Date","Abundance", fill='Family')  +ylab("Relative Bacterial Abundance (> 1%)")+facet_grid(Treatment ~ .)#+scale_fill_manual(values=cbPalette)
TrtDateFamilyplot+ theme(axis.text.x = element_text(angle = 45, hjust = 1))#+ theme(legend.justification=c(0.05,0.95), legend.position=c(0.05,0.95))

```



```{r}

```
#Reference vs Modified (Heavy and Moderate combined)
##Observed Species

```{r,echo=FALSE}
plot_richness(physeq, x="Date",color="Date", measures=c("Observed"),)+geom_boxplot(aes(x=Date, y=value, color=Date), alpha=0.05)+facet_wrap(~Treatment2)+ylab("Observed Species")

```

##Simpson Diversity
```{r,echo=FALSE}

plot_richness(physeq, x="Date",color="Date", measures=c("Simpson"),)+geom_boxplot(aes(x=Date, y=value, color=Date), alpha=0.05)+facet_wrap(~Treatment2)+ylab("Simpson Diversity")

```



#Taxa Plots (Modified vs Reference)
```{r filteringForFamily ,echo=FALSE}
GPr  = transform_sample_counts(physeq, function(x) x / sum(x) ) #transform samples based on relative abundance
GPrPhylum=tax_glom(GPr, "Phylum")
PhylumLevel = filter_taxa(GPrPhylum, function(x) mean(x) > 1e-3, TRUE) #filter out any taxa lower tha 0.1%
GPrFamily=tax_glom(GPr,"Family")
FamilyLevel = filter_taxa(GPrFamily, function(x) mean(x) > 1e-2, TRUE) #filter out any taxa lower tha 1%

```

##Phylum Level Relative Abundance
```{r Summarizing3, echo=FALSE}
df <- psmelt(PhylumLevel)
df$Abundance=df$Abundance*100
Trtdata <- ddply(df, c("Phylum", "Treatment2"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
```
```{r TreatmentPlot3, echo=FALSE}
Plot=ggplot(Trtdata, aes(x=Treatment2, y=mean,fill=Treatment2))+facet_wrap(~Phylum)
Plot+geom_bar(stat="identity")+ geom_errorbar(aes(ymin=mean-se, ymax=mean+se),col="black")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ylab("Relative Abundance (> 0.1%)")
```

##Family Level Relative Abundance
```{r Summarizing, echo=FALSE}
df <- psmelt(FamilyLevel)
df$Abundance=df$Abundance*100
Trtdata <- ddply(df, c("Family", "Treatment2"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
```
```{r TreatmentPlot, echo=FALSE}
Plot=ggplot(Trtdata, aes(x=Treatment2, y=mean,fill=Treatment2))+facet_wrap(~Family)
Plot+geom_bar(stat="identity")+ geom_errorbar(aes(ymin=mean-se, ymax=mean+se),col="black")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ylab("Relative Abundance (> 1%)")
```

#Betadisper
##By Treatment Only
### Permutest
``` {r, echo=FALSE}
GPdist=phyloseq::distance(physeq, "jaccard")
beta <- betadisper(GPdist, sample_data(physeq)$Treatment)
permutest(beta, pairwise= FALSE, permutations= 999)
```

###Tukey HS
```{r, echo=FALSE}
beta.HSD=TukeyHSD(beta)
beta.HSD$group
```

###Betadisper plot
I can make the plots in vegan look better later( Black=Heavy, Red= Moderate, Green= Reference ). Confidence intervals are 95% CI
```{r, echo=FALSE}
plot(beta, label=FALSE,ellipse=TRUE, hull=FALSE, conf=0.95, segments = FALSE)
```

### Boxplot of distance from centroid
```{r, echo= FALSE}
boxplot(beta)
```

##Betadisper by Treatment x Date
### Permutest
``` {r, echo=FALSE}
GPdist=phyloseq::distance(physeq, "jaccard")
beta <- betadisper(GPdist, sample_data(physeq)$DateTreat)
permutest(beta, pairwise= FALSE, permutations= 999)
```

```

###Betadisper plot
I can make the plots in vegan look better later. Confidence intervals are 95% CI
```{r, echo=FALSE}
plot(beta, label=FALSE,ellipse=TRUE, hull=FALSE, conf=0.95, segments = FALSE)
```

### Boxplot of distance from centroid
```{r, echo= FALSE}
boxplot(beta)
```
